---
title: "Preprocess Scripts"
date: "2024-07-10"
---

All the filepaths are relative to the .Rproj file.

```{r setup, include=FALSE}
#Before we start, set the working directory to where the .Rproj lives. here::here()
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include = FALSE, echo = FALSE, eval = FALSE}
# This better output as your .Rproj location. NOT the .Rmd location.
getwd()
```

Let us preprocess the data.

# 1 Nanopore Preprocess

Find your files to the three different .bedmethyl files:

        "Su567-BTRC_131_Control_F_1.wf_mods.bedmethyl.gz"         
        "Su567-BTRC_131_Control_F_2.wf_mods.bedmethyl.gz" 
        "Su567-BTRC_131_Control_F_ungrouped.wf_mods.bedmethyl.gz"                                            


Su349_BTRC_51_Control_M_analysis  Su636_BTRC_168_Control_M_merge_analysis
Su395_BTRC_69_Control_M_analysis  Su779_BTRC_266_AUD_M_analysis
[stevbroo@i68 Nanopore]$ realpath .
/N/project/ICMH/CCBB/Yun

```{r}
source('src/preprocess/1_combine_haplotagged_bedmethyl.R')

### Process samples
#### File paths needed:
nano_files_path <- 'data/Dec2024_comparison/Nanopore'

sample_51_dir <- file.path(nano_files_path, 'Su349_BTRC_51_Control_M_analysis')
sample_69_dir <- file.path(nano_files_path, 'Su395_BTRC_69_Control_M_analysis')
sample_168_dir <- file.path(nano_files_path, 'Su636_BTRC_168_Control_M_merge_analysis')
sample_266_dir <- file.path(nano_files_path, 'Su779_BTRC_266_AUD_M_analysis')


nano_files_path_epic <- '../nanopore_epic/data'
sample_541_dir <- file.path(nano_files_path_epic, 'sample_4059246541_analysis')
sample_577_dir <- file.path(nano_files_path_epic, 'sample_4059246577_analysis')
sample_475_dir <- file.path(nano_files_path_epic, 'sample_4059249475_Analysis')
sample_500_dir <- file.path(nano_files_path_epic, 'sample_4059246500_analysis')


### Output directory
nano_data_dir <- '../../data/nanopore/'


```

Merge all of the haplotypes together and save as an .rds file in the output directory.






```{r, eval = FALSE}
######### EM-Seq sample pairs
# 51
mergeHapsAllMethylation(bedmethyl_file_hap1 = file.path(sample_51_dir, get_bedmethyl_files(sample_51_dir)['h1']),
                        bedmethyl_file_hap2 = file.path(sample_51_dir, get_bedmethyl_files(sample_51_dir)['h2']),
                        bedmethyl_file_ungrouped = file.path(sample_51_dir, get_bedmethyl_files(sample_51_dir)['ungr']),
                        filename_out = file.path(nano_data_dir,
                                                  get_bedmethyl_files(sample_51_dir)['sample_name']))

gc()

# 69
mergeHapsAllMethylation(bedmethyl_file_hap1 = file.path(sample_69_dir, get_bedmethyl_files(sample_69_dir)['h1']),
                        bedmethyl_file_hap2 = file.path(sample_69_dir, get_bedmethyl_files(sample_69_dir)['h2']),
                        bedmethyl_file_ungrouped = file.path(sample_69_dir, get_bedmethyl_files(sample_69_dir)['ungr']),
                        filename_out = file.path(nano_data_dir,
                                                 get_bedmethyl_files(sample_69_dir)['sample_name']))

gc()
# 168
mergeHapsAllMethylation(bedmethyl_file_hap1 = file.path(sample_168_dir, get_bedmethyl_files(sample_168_dir)['h1']),
                        bedmethyl_file_hap2 = file.path(sample_168_dir, get_bedmethyl_files(sample_168_dir)['h2']),
                        bedmethyl_file_ungrouped = file.path(sample_168_dir, get_bedmethyl_files(sample_168_dir)['ungr']),
                        filename_out = file.path(nano_data_dir,
                                                 get_bedmethyl_files(sample_168_dir)['sample_name']))

gc()

# 266
mergeHapsAllMethylation(bedmethyl_file_hap1 = file.path(sample_266_dir, get_bedmethyl_files(sample_266_dir)['h1']),
                        bedmethyl_file_hap2 = file.path(sample_266_dir, get_bedmethyl_files(sample_266_dir)['h2']),
                        bedmethyl_file_ungrouped = file.path(sample_266_dir, get_bedmethyl_files(sample_266_dir)['ungr']),
                        filename_out = file.path(nano_data_dir,
                                                 get_bedmethyl_files(sample_266_dir)['sample_name']))
gc()
```



```{r}
# 475
mergeHapsAllMethylation(bedmethyl_file_hap1 = file.path(sample_475_dir, get_bedmethyl_files(sample_475_dir)['h1']),
                        bedmethyl_file_hap2 = file.path(sample_475_dir, get_bedmethyl_files(sample_475_dir)['h2']),
                        bedmethyl_file_ungrouped = file.path(sample_475_dir, get_bedmethyl_files(sample_475_dir)['ungr']),
                        filename_out = file.path(nano_data_dir, 'nano_475.rds'))

gc()

mergeHapsAllMethylation(bedmethyl_file_hap1 = file.path(sample_500_dir, get_bedmethyl_files(sample_500_dir)['h1']),
                        bedmethyl_file_hap2 = file.path(sample_500_dir, get_bedmethyl_files(sample_500_dir)['h2']),
                        bedmethyl_file_ungrouped = file.path(sample_500_dir, get_bedmethyl_files(sample_500_dir)['ungr']),
                        filename_out = file.path(nano_data_dir, 'nano_500.rds'))

gc()
# 541
mergeHapsAllMethylation(bedmethyl_file_hap1 = file.path(sample_541_dir, get_bedmethyl_files(sample_541_dir)['h1']),
                        bedmethyl_file_hap2 = file.path(sample_541_dir, get_bedmethyl_files(sample_541_dir)['h2']),
                        bedmethyl_file_ungrouped = file.path(sample_541_dir, get_bedmethyl_files(sample_541_dir)['ungr']),
                        filename_out = file.path(nano_data_dir, 'nano_541.rds'))

gc()
# 577
mergeHapsAllMethylation(bedmethyl_file_hap1 = file.path(sample_577_dir, get_bedmethyl_files(sample_577_dir)['h1']),
                        bedmethyl_file_hap2 = file.path(sample_577_dir, get_bedmethyl_files(sample_577_dir)['h2']),
                        bedmethyl_file_ungrouped = file.path(sample_577_dir, get_bedmethyl_files(sample_577_dir)['ungr']),
                        filename_out = file.path(nano_data_dir, 'nano_577.rds'))
gc()
```


#################################################################

# EM-Seq Processing

As above, start with finding your input filepaths.

BTRC_168_BTRC_168_Control.CX_report.txt.gz
BTRC_266_BTRC_266_AUD.CX_report.txt.gz


```{r}
source('2_save_EM_rds.R')


########## Pathing
em_input_dir <- '../../data/Dec2024_comparison/EM'

#input
em_51_path <- file.path(em_input_dir,
                        'BTRC_51_BTRC_51_Control.CX_report.txt.gz')

em_69_path <- file.path(em_input_dir,
                         'BTRC_69_BTRC_69_Control.CX_report.txt.gz')

em_168_path <- file.path(em_input_dir,
                         'BTRC_168_BTRC_168_Control.CX_report.txt.gz')

em_266_path <- file.path(em_input_dir,
                         'BTRC_266_BTRC_266_AUD.CX_report.txt.gz')

### Output directory
em_data_dir <- '../../data/em-seq'
```

```{r,eval = FALSE}
em_proc(em_filepath = em_51_path, output_filepath = file.path(em_data_dir, 'em_51.rds'))
gc()
em_proc(em_filepath = em_69_path, output_filepath = file.path(em_data_dir, 'em_69.rds'))
gc()
em_proc(em_filepath = em_168_path, output_filepath = file.path(em_data_dir, 'em_168.rds'))
gc()
em_proc(em_filepath = em_266_path, output_filepath = file.path(em_data_dir, 'em_266.rds'))
gc()
```

#################################################################

# DMR calling on one nanopore sample

```{r}
source('3_make_nano_DMR.R')


# pathing
sample_dir <- "../nanopore_epic/data/sample_4059249500_analysis"
dmrs_out_dir <- "data/dmrs"


dmrs_pipeline(sample_dir, dmrs_out_dir, "500", min_cov = 10)
gc()
```

#################################################################

# Creating annotations for gene regions, TSS, and CpG islands

```{r}
source('src/preprocess/4_create_annotations.R')
make_annotations(reference_directory = 'data/reference')
```


